---
title: "Population Transportability in Network Meta-Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Population Transportability in Network Meta-Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nmatransport)
```

## Introduction

Network Meta-Analysis (NMA) assumes transitivity: the relative effect of A vs B can be estimated from A vs C and B vs C. This assumption is violated if effect modifiers (e.g., patient age, severity) differ systematically across the comparisons in the network.

The `nmatransport` package provides tools to:
1. Assess the transitivity assumption by quantifying covariate overlap.
2. Reweight the network using Entropy Balancing to "transport" the results to a specific target population.

## Data Structure

We start with a simulated NMA dataset containing treatment effects and study-level covariates.

```{r data}
data(nma_data)
head(nma_data)
```

## Step 1: Transitivity Assessment

We check if effect modifiers (`age_mean`, `bmi_mean`) are balanced across the different edges of the network.

```{r transitivity}
trans_results <- assess_transitivity(nma_data, modifiers = c("age_mean", "bmi_mean"))
print(trans_results$global_transitivity_violation)
```

A high transitivity violation score suggests that the standard NMA assumptions may not hold, necessitating population transportability methods.

## Step 2: Transporting Evidence

Suppose a clinician wants to know the treatment rankings for an older population with a higher BMI. We define the target population profile:

```{r target}
target_pop <- list(age_mean = 65, bmi_mean = 30)
```

We compute transport weights for each study so that the weighted average of the network matches the target population:

```{r weights}
weights <- compute_nma_weights(nma_data, target_pop)
head(weights)
```

## Step 3: Run Transported NMA

Finally, we run the NMA using the computed weights. This down-weights studies that are dissimilar to our target population and up-weights studies that closely match it.

```{r transported_nma}
if (requireNamespace("netmeta", quietly = TRUE)) {
  # Unweighted standard NMA
  res_standard <- transported_nma(nma_data)
  print(res_standard)
  
  # Transported NMA
  res_transported <- transported_nma(nma_data, weights = weights)
  print(res_transported)
}
```

By comparing the treatment rankings (P-scores), we can observe how the optimal treatment changes depending on the specific patient population.
